Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi

### 2013-08-01 writeup/hat-vldb.tex
Last edits

@@

\title{Highly Available Transactions: Virtues and Limitations\\(Extended Version)}
{\author{Peter Bailis{\fontsize{12}{14}$^\dagger$}, Aaron Davidson{\fontsize{12}{14}$^\dagger$}, Alan Fekete{\fontsize{12}{14}$^\diamond$}, Ali Ghodsi{\fontsize{12}{14}$^{\dagger,\ddagger}$}, Joseph M. Hellerstein{\fontsize{12}{14}$^\dagger$}, Ion Stoica{\fontsize{12}{14}$^\dagger$}\\{\affaddr{\fontsize{12}{14}$^\dagger$}\hspace{.5mm}UC Berkeley\hspace{4mm}{\fontsize{12}{14}$^\diamond$}\hspace{.5mm}University of Sydney\hspace{4mm}{\fontsize{12}{14}$^\ddagger$}\hspace{.5mm}KTH/Royal Institute of Technology}}}
\maketitle

@@

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
\pagebreak

\pagebreak
\begin{appendix}

\input{hat-appendix.tex}
\end{appendix}
\end{document}

>>>>

\newif\iftechreport
%\techreporttrue
\techreportfalse

@@

\title{Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi}
{\author{Peter Bailis{\fontsize{12}{14}$^\dagger$}, Aaron Davidson{\fontsize{12}{14}$^\dagger$}, Alan Fekete{\fontsize{12}{14}$^\diamond$}, Ali Ghodsi{\fontsize{12}{14}$^{\dagger,\ddagger}$}, Joseph M. Hellerstein{\fontsize{12}{14}$^\dagger$}, Ion Stoica{\fontsize{12}{14}$^\dagger$}\\{\affaddr{\fontsize{12}{14}$^\dagger$}\hspace{.5mm}UC Berkeley\hspace{4mm}{\fontsize{12}{14}$^\diamond$}\hspace{.5mm}University of Sydney\hspace{4mm}{\fontsize{12}{14}$^\ddagger$}\hspace{.5mm}KTH/Royal Institute of Technology}}}
\maketitle

@@

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page

\iftechreport
\pagebreak
\begin{appendix}

\input{hat-appendix.tex}
\end{appendix}
\fi
\end{document}

### 2013-08-01 writeup/model.tex
Last edits

Traditionally, a system provides {\textbf{high availability}} if every
user that can contact a correct (non-failing) server eventually
receives a response from that server, even in the presence of
arbitrary, indefinitely long network partitions between
servers~\cite{gilbert-cap}.\footnote{Under this definition from the
  distributed systems literature, systems that require a majority of
  servers to be online are not available. Similarly, a system which
  ensures that clients receive a response with high (but not perfect)
  probability is not available. This is a stringent requirement but it
  matches the assumptions made in the CAP Theorem~\cite{gilbert-cap}
  and guarantees low latency~\cite{abadi-pacelc}.} As in a standard
distributed database, designated servers might perform operations for
different data items; a server that can handle an operation for a
given data item is called a \textit{replica} for that
item.\footnote{There is a further distinction between a \textit{fully
    replicated} system, in which all servers are replicas for all data
  items, and a \textit{partially replicated} system, in which at least
  one server acts as a replica for only a subset of all data
  items. For generality, and, for applicability to many modern
  ``sharded'' or ``partitioned'' data storage systems~\cite{ bigtable,
    pnuts, spanner, dynamo, hstore}, we consider partially replicated
  systems.}

>>>>

Traditionally, a system provides {\textbf{high availability}} if every
user that can contact a correct (non-failing) server eventually
receives a response from that server, even in the presence of
arbitrary, indefinitely long network partitions between
servers~\cite{gilbert-cap}.\footnote{Under this definition from the
  distributed systems literature, systems that require a majority of
  servers to be online are not available. Similarly, a system which
  guarantees that servers provide a response with high probability is
  not available. This admittedly stringent requirement matches the
  assumptions made in the CAP Theorem~\cite{gilbert-cap} and
  guarantees low latency~\cite{abadi-pacelc}.} As in a standard
distributed database, designated servers might perform operations for
different data items; a server that can handle an operation for a
given data item is called a \textit{replica} for that
item.\footnote{There is a further distinction between a \textit{fully
    replicated} system, in which all servers are replicas for all data
  items, and a \textit{partially replicated} system, in which at least
  one server acts as a replica for only a subset of all data
  items. For generality, and, for applicability to many modern
  ``sharded'' or ``partitioned'' data storage systems~\cite{ bigtable,
    pnuts, spanner, dynamo, hstore}, we consider partially replicated
  systems.}

### 2013-08-01 writeup/relatedwork.tex
Last edits

There has been a recent resurgence of interest in distributed
multi-object semantics, both in academia~\cite{kraska-s3, gstore,
  eiger, walter,calvin, swift} and industry~\cite{orleans,spanner}. As
discussed in Section~\ref{sec:modernacid}, classic ACID databases
provide strong semantics but their lock-based and traditional
multi-versioned implementations are unavailable in the presence of
partitions~\cite{bernstein-book, gray-isolation}. Notably, Google's
Spanner provides strong one-copy serializable transactions. While
Spanner is highly specialized for Google's read-heavy workload, it
relies on two-phase commit and two-phase locking for read/write
transactions~\cite{spanner}. As we have discussed, the latency
penalties associated with this design choice are fundamental to
serializability. For users willing to tolerate these costs, Spanner,
or similar strongly consistent, unavailable systems---including
Calvin~\cite{calvin}, G-Store~\cite{gstore}, HBase,
HStore~\cite{hstore}, Orleans~\cite{orleans}, Postgres-R~\cite{kemme-thesis},
Walter~\cite{walter}, and a range of snapshot isolation
techniques~\cite{daudjee-session, daudjee-snapshot}---are a reasonable
choice.

With HATs, we seek an alternative set of transactional semantics that
are still useful but do not violate requirements for high availability
or low latency. Recent systems proposals such as Swift~\cite{swift},
Eiger~\cite{eiger}, and Bolt-on Causal Consistency~\cite{bolton}
provide transactional causal consistency guarantees with varying
availability and represent a new class of sticky HAT systems. There
are infinitely many HAT models (i.e., always reading value 1 is
incomparable with always returning value 2), but a recent report from
UT Austin shows that no model stronger than causal consistency can be
achieved in a sticky highly available, \textit{one-way convergent}
system~\cite{cac}. This result is promising and complementary to our
results for general-purpose convergent data stores. Finally, Burkhardt
et al. have concurrently developed an axiomatic specification for
eventual consistency; their work-in-progress report contains alternate
formalism for several HAT guarantees~\cite{burkhardt-txns}.

>>>>

There has been a recent resurgence of interest in distributed
multi-object semantics, both in academia~\cite{kraska-s3, gstore,
  eiger, walter,calvin, swift} and industry~\cite{orleans,spanner}. As
discussed in Section~\ref{sec:modernacid}, classic ACID databases
provide strong semantics but their lock-based and traditional
multi-versioned implementations are unavailable in the presence of
partitions~\cite{bernstein-book, gray-isolation}. Notably, Google's
Spanner provides strong one-copy serializable transactions. While
Spanner is highly specialized for Google's read-heavy workload, it
relies on two-phase commit and two-phase locking for read/write
transactions~\cite{spanner}. As we have discussed, the latency
penalties associated with this design choice are fundamental to
serializability. For users willing to tolerate these costs, Spanner,
or similar strongly consistent, unavailable
systems~\cite{kemme-classification}---including Calvin~\cite{calvin},
G-Store~\cite{gstore}, HBase, HStore~\cite{hstore},
Orleans~\cite{orleans}, Postgres-R~\cite{kemme-thesis},
Walter~\cite{walter}, and a range of snapshot isolation
techniques~\cite{daudjee-session, daudjee-snapshot}---are a reasonable
choice.

With HATs, we seek an alternative set of transactional semantics that
are still useful but do not violate requirements for high availability
or low latency. Recent systems proposals such as Swift~\cite{swift},
Eiger~\cite{eiger}, and Bolt-on Causal Consistency~\cite{bolton}
provide transactional causal consistency guarantees with varying
availability and represent a new class of sticky HAT systems. There
are infinitely many HAT models (i.e., always reading value 1 is
incomparable with always returning value 2), but a recent report from
UT Austin shows that no model stronger than causal consistency is
achievable in a sticky highly available, \textit{one-way convergent}
system~\cite{cac}. This result is promising and complementary to our
results for general-purpose convergent data stores. Finally, Burkhardt
et al. have concurrently developed an axiomatic specification for
eventual consistency; their work-in-progress report contains alternate
formalism for several HAT guarantees~\cite{burkhardt-txns}.
